### 2021-04-19 paper/ase21/analyzer.tex
[ASE21] Updated analyzer

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{\textwidth}
    \[
      \footnotesize
      \begin{array}{r@{~}c@{~}ll}
        \asemi{\kwlet \; \x = \expr}
        (\lab, \tys)(\aelem) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \aenv_\x \}, \emp) &
        \text{where} \;
        \aenv_\x = \aenv[\x \mapsto \aty] \wedge
        \aty = \aseme{\expr}(\aenv)\\

        \asemi{\x = \kwrl \expr_0 \; \expr_1 \cdots \expr_n \kwrr}
        (\lab, \tys)(\aelem) &=&
        \left(
          \begin{array}{@{}l@{}}
            \{ (\lab_\func, \tys_a) \mapsto \aenv_{\func, \tys_a} \mid \func \in
              F \wedge \tys_a \in T_a
            \},\\
            \{ (\func, \tys_a) \mapsto \{ (\getnext(\lab), \tys, \x) \} \mid
              \func \in F \wedge \tys_a \in T_a
            \}\\
          \end{array}
        \right) &
        \text{where} \;
        \left\{
          \begin{array}{@{}l@{}}
            \aty_i = \aseme{\expr_i}(\aenv) \wedge T_a = \{ \upcasts([ \ty_1,
            \cdots, \ty_n ] \mid \ty_i \in \aty_i \} \wedge\\
            F = \{ \func \in \aty_0 \mid \func = \kwdef \; \_ (\p_1, \cdots,
            \kwsl \cdots, \p_m \kwsr). \; \lab_\func \} \wedge\\
            \aenv_{\func, \tys_a} = \aenv[\p_1 \mapsto \{ \tys_a[1] \}, \cdots,
            \p_m \mapsto \{ \tys_a[m] \}]\\
          \end{array}
        \right.\\

        \asemi{\kwassert \; \expr}
        (\lab, \tys)(\aelem) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \aenv_\mt \}, \emp) &
        \text{where} \; \aenv_\mt = \refine(\expr, \true)(\aenv)\\

        \asemi{\kwif \; \expr \; \lab_\mt \; \lab_\mf}
        (\lab, \tys)(\aelem) &=&
        (\{ (\lab_\mt, \tys) \mapsto \aenv_\mt, (\lab_\mf, \tys) \mapsto
        \aenv_\mf \}, \emp) &
        \text{where} \;
        \aenv_\mt = \refine(\expr, \true)(\aenv) \wedge
        \aenv_\mf = \refine(\expr, \false)(\aenv)\\

        \asemi{\kwreturn \; \expr}
        (\lab, \tys)(\aelem) &=&
        (\{ (\lab_r, \tys_r) \mapsto \aenv_r \mid (\lab_r, \tys_r, \x) \in
        \retp(\getfunc(\lab), \tys)\}, \emp) &
        \text{where} \;
        \aenv_r = \rmap(\lab_r, \tys_r)[\x \mapsto \aty] \wedge
        \aty = \aseme{\expr}(\aenv)\\

        \asemi{\x = \expr}
        (\lab, \tys)(\aelem) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \aenv_\x \}, \emp) &
        \text{where} \;
        \aenv_\x = \aenv[\x \mapsto \aty] \wedge
        \aty = \aseme{\expr}(\aenv)\\

        \asemi{\refer \kwsl \expr_0 \kwsr = \expr_1}
        (\lab, \tys)(\aelem) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \aenv \}, \emp)
      \end{array}
    \]
    \caption{Instructions with $\asemi{\inst}: (\labset \times \tyset^*)
    \rightarrow \adom \rightarrow \adom$ where $\aelem = (\rmap, \retp) \wedge
    \aenv = \aelem(\lab, \tys)$.}
  \end{subfigure}
  \begin{subfigure}[b]{0.48\textwidth}
    \[
      \footnotesize
      \asemr{\x}(\aenv) = \aenv(\x)
      \qquad
      \asemr{\refer \kwsl \expr \kwsr}(\aenv) = \{ \ty[\str] \mid \ty \in
      \asemr{\refer}(\aenv) \wedge \str \in \aseme{\expr}(\aenv) \}
    \]
    \caption{References with $\asemr{\refer}: \aenvset \rightarrow \atyset$.}
  \end{subfigure}
  \begin{subfigure}[b]{\textwidth}
    \[
      \footnotesize
      \begin{array}{r@{~}c@{~}lr@{~}c@{~}lr@{~}c@{~}l}
        \aseme{\tname \; \kwcl \fields \kwcr}(\aenv)
        &=& \multicolumn{7}{@{}l}{
          \left\{
            \begin{array}{ll}
              \{ \tnormal(\ty) \mid \ty \in \aseme{\fields(\code{Value})}(\aenv)
              \} & \text{if} \; \tname = \text{Completion} \wedge
              \fields(\code{Type}) = \nconst{normal}\\

              \{ \tabrupt \} & \text{if} \; \tname = \text{Completion} \wedge
              \fields(\code{Type}) \neq \nconst{normal}\\

              \{ \tname \} & \text{otherwise}\\
            \end{array}
          \right.
          \quad \aseme{\prim}(\aenv) =
          \left\{
            \begin{array}{ll}
              \{ \tnum \} & \text{if} \; \prim = \num\\
              \{ \tbigint \} & \text{if} \; \prim = \bigint\\
              \{ \tsymb \} & \text{if} \; \prim = \symb\\
              \{ \prim \} & \text{otherwise}\\
            \end{array}
          \right.
        }

        \\

        \aseme{\const}(\aenv) &=& \const &

        \aseme{\clist{}}(\aenv) &=& \clist{} &

        \aseme{\clist{\expr_0, \cdots, \expr_n}}(\aenv)
        &=& \{ \clist{\ty} \mid \ty \in \bigjoin_{0 \leq i
        \leq n} \aseme{\expr_i}(\aenv) \}

        \\

        \aseme{\expr: \ty}(\aenv)
        &=& \{ \ty' \subtype \ty \mid \ty' \in
        \aseme{\expr}(\aenv) \} &

        \aseme{\refer \kwexists}(\aenv)
        &=& \{ \ty \neq \tabsent \mid \ty \in
        \aseme{\expr}(\aenv) \} &

        \aseme{\refer}(\aenv)
        &=& \asemr{\refer}(\aenv) \setminus \{ \tabsent \}

        \\

        \aseme{\uop \; \expr}(\aenv)
        &=& \{ \abs{\uop} \ty \mid \ty \in
        \aseme{\expr}(\aenv) \} &

        \aseme{\expr_0 \bop \expr_1}(\aenv)
        &=& \multicolumn{4}{@{}l}{\{ \ty_0 \abs{\bop} \ty_1 \mid \ty_0 \in
        \aseme{\expr_0}(\aenv) \wedge \ty_1 \in \aseme{\expr_1}(\aenv) \}}

        \\
      \end{array}
    \]
    \caption{Expressions with $\aseme{\expr}: \aenvset \rightarrow \atyset$.}
  \end{subfigure}
  \caption{Abstract semantics of the modified $\ires$ for a program $\prog =
  (\getfunc, \getinst, \getnext)$.}
  \vspace*{-1.5em}
  \label{fig:abs-sem}
\end{figure*}

@@

Based on abstract states, we define the abstract semantics of instructions,
references, and expressions in the modified $\ires$ described in
Figure~\ref{fig:abs-sem}.  To prevent the explosion of the number of type-sensitive
views, we upcast the type of each argument before function calls with the
following function:
\begin{figure}[H]
  \centering
  \vspace{-0.5em}
  \resizebox{0.8\columnwidth}{!}{$
    \upcast(\ty) = \left\{
      \begin{array}{ll}
        \tnormal(\upcast(\ty')) & \text{if} \; \ty = \tnormal(\ty')\\
        \clist{\upcast(\ty')} & \text{if} \; \ty = \clist{\ty'}\\
        \tstr & \text{if} \; \ty = \str\\
        \tbool & \text{if} \; \ty = \bool\\
        \ty & \text{otherwise}\\
      \end{array}
    \right.
  $}
  \vspace{-0.5em}
\end{figure} \noindent
and $\upcasts$ denotes a point-wise extension of $\upcast$ for type sequences.
For branches and assertions, we apply the $\refine$ function that prunes out
infeasible parts in abstract environments based on their conditions to increase
analysis precision.  We will explain its detail in Section~\ref{sec:refine}.
Then, we define the abstract semantics $\asem{\prog}$ of a program $\prog =
(\getfunc, \getinst, \getnext)$ as the least fixpoint of abstract transfer
$\atransfer: \adom \rightarrow \adom$:
\begin{figure}[H]
  \centering
  \vspace{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}l}
      \asem{\prog} &=& \lim_{n \rightarrow \infty}(\atransfer)^n(\iaelem)\\
      \atransfer(\aelem) &=& \aelem \join \left(
        \bigjoin_{(\lab, \tys) \in \domain{\rmap}}
        \asemi{\getinst(\lab)}(\lab, \tys)(\aelem)
      \right)\\
    \end{array}
  $}
  \vspace{-0.5em}
\end{figure} \noindent
where $\aelem = (\rmap, \_)$ and $\iaelem$ denotes the initial abstract state.
As described in Section~\ref{sec:overview}, $\iaelem$ contains entry points of
all syntax-directed algorithms having no additional parameters and built-in
algorithms with appropriate abstract environments.  For syntax-directed
algorithms, we construct its abstract environment containing the variable
$\this$ with a production type and other variables for non-terminals.  For
example, the upper-left syntax-directed algorithm in Figure~\ref{fig:example} is
initialized with the following abstract environment:
\begin{figure}[H]
  \centering
  \vspace{-0.5em}
  \resizebox{\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}l}
      \this &\mapsto& \{ \code{AssignmentExpression} \},\\
      \code{LeftHandSideExpression} &\mapsto& \{ \code{LeftHandSideExpression} \},\\
      \code{AssignmentExpression} &\mapsto& \{ \code{AssignmentExpression} \}
    \end{array}
  $}
  \vspace{-0.5em}
\end{figure} \noindent
For built-in algorithms, we assign pre-defined variables $\this$, $\args$, and
$\NewTarget$ with corresponding types and parameters with $\tjs$ types.  For
example, the following abstract environment is for the built-in
algorithm $\jscode{Math.round}$ at the upper-right in Figure~\ref{fig:example}:
\begin{figure}[H]
  \centering
  \vspace{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}lr@{~}c@{~}l}
      \this &\mapsto& \{ \tjs \}, &
      \args &\mapsto& \{ \clist{\tjs} \},\\
      \NewTarget &\mapsto& \{ \code{Object}, \undefval \}, &
      \x &\mapsto& \tjs\\
    \end{array}
  $}
  \vspace{-0.5em}
\end{figure}

>>>>

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.52\textwidth}
    \resizebox{\textwidth}{!}{$
      \begin{array}{@{}l@{~}c@{~}l}
        \asemi{\kwlet \; \x = \expr}
        (\lab, \tys)(\aelem) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \aenv[\x \mapsto
        \aseme{\expr}(\aenv)] \}, \emp)\\

        \asemi{\x = \kwrl \expr \; \expr_1 \cdots \expr_n \kwrr}
        (\lab, \tys)(\aelem) &=& (\rmap', \retp')\\
        \multicolumn{3}{@{}l}{
          \text{where} \;
          \left\{
            \begin{array}{l@{}}
              \aty = \aseme{\expr}(\aenv) \;\wedge\;
              \aty_1 = \aseme{\expr_1}(\aenv) \;\wedge\;
              \cdots \;\wedge\;
              \aty_n = \aseme{\expr_n}(\aenv) \;\wedge\;\\

              T' = \{ \upcasts([ \ty_1, \cdots, \ty_n ]) \mid \ty_1 \in \aty_1
              \;\wedge\; \cdots \;\wedge\; \ty_n \in \aty_n \} \;\wedge\;\\

              \func = \kwdef \; \f (\p_1, \cdots, \kwsl \cdots, \p_{m_\func}
              \kwsr). \; \lab_\func \;\wedge\;\\

              \aenv_{\func, \tys'} = [\p_1 \mapsto \{ \tys'[1] \}, \cdots,
              \p_{m_\func} \mapsto \{ \tys'[m] \}] \;\wedge\;\\

              \rmap' = \{ (\lab_\func, \tys') \mapsto \aenv_{\func, \tys'}
              \mid \func \in \aty \;\wedge\; \tys' \in T' \} \;\wedge\;\\

              \retp' = \{ (\func, \tys') \mapsto \{ (\getnext(\lab), \tys, \x)
              \} \mid \func \in \aty \;\wedge\; \tys' \in T' \}\\
            \end{array}
          \right.
        }\\

        \asemi{\kwreturn \; \expr}
        (\lab, \tys)(\aelem) &=&
        (\{ (\lab_r, \tys_r) \mapsto \aenv_r \mid
        (\lab_r, \tys_r, \x) \in R\}, \emp)\\
        \multicolumn{3}{@{}l}{
          \text{where} \; R = \retp(\getfunc(\lab), \tys) \;\wedge\; \aenv_r =
          \rmap(\lab_r, \tys_r)[\x \mapsto \aseme{\expr}(\aenv)]
        }\\

        \asemi{\kwassert \; \expr}
        (\lab, \tys)(\aelem) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \pass(\expr, \true)(\aenv) \}, \emp)\\

        \asemi{\kwif \; \expr \; \lab_\mt \; \lab_\mf}
        (\lab, \tys)(\aelem) &=&
        (\{ (\lab_\mt, \tys) \mapsto \pass(\expr, \true)(\aenv),\\ &&
        \phantom{(\{} (\lab_\mf, \tys) \mapsto \pass(\expr, \false)(\aenv) \}, \emp)\\

        \asemi{\x = \expr}
        (\lab, \tys)(\aelem) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \aenv[\x \mapsto
        \aseme{\expr}(\aenv)] \}, \emp)\\

        \asemi{\refer \kwsl \expr_0 \kwsr = \expr_1}
        (\lab, \tys)(\aelem) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \aenv \}, \emp)\\

        \multicolumn{3}{@{}l}{\text{where} \; \aelem = (\rmap, \retp) \;\wedge\;
        \aenv = \rmap(\lab, \tys)}
      \end{array}
    $}
    \caption{Instructions with $\asemi{\inst}: (\labset \times \tyset^*)
    \rightarrow \adom \rightarrow \adom$.}
  \end{subfigure}
  \begin{minipage}[b][0.3\textheight][s]{0.46\textwidth}
    \begin{subfigure}[b]{\textwidth}
      \centering
      \vspace*{0.1em}
      \resizebox{0.8\textwidth}{!}{$
        \begin{array}{l@{~}c@{~}l}
          \asemr{\x}(\aenv) &=& \aenv(\x)\\
          \asemr{\refer \kwsl \expr \kwsr}(\aenv) &=& \{ \ty[\str] \mid \ty \in
          \asemr{\refer}(\aenv) \wedge \str \in \aseme{\expr}(\aenv) \}\\
        \end{array}
      $}
      \caption{References with $\asemr{\refer}: \aenvset \rightarrow \atyset$.}
      \vspace*{0.5em}
    \end{subfigure}
    \begin{subfigure}[b]{\textwidth}
      \centering
      \resizebox{\textwidth}{!}{$
        \begin{array}{l@{~}c@{~}l@{}}
          \multicolumn{3}{l}{
            \aseme{\num}(\aenv) = \tnum \qquad
            \aseme{\bigint}(\aenv) = \tbigint \qquad
            \aseme{\symb}(\aenv) = \tsymb
          }\\

          \multicolumn{3}{l}{
            \aseme{\prim}(\aenv) = \prim \qquad
            \aseme{\const}(\aenv) = \const \qquad
            \aseme{\clist{}}(\aenv) = \clist{}
          }\\

          \multicolumn{3}{l}{
            \aseme{\clist{\expr_0, \cdots, \expr_n}}(\aenv) = \{ \clist{\ty} \mid
            \ty \in \bigjoin_{0 \leq i \leq n} \aseme{\expr_i}(\aenv) \}
          }\\

          \multicolumn{3}{l}{
            \aseme{\code{Completion} \; \kwcl \cdots, \code{Type}:
            \expr_0, \code{Value}: \expr_1, \cdots \kwcr}(\aenv) =
          }\\
          \multicolumn{3}{l}{
            \qquad \left\{
              \begin{array}{ll}
                \{ \tnormal(\ty) \mid \ty \in \aseme{\expr_1}(\aenv) \} &
                \text{if} \; \aseme{\expr_0} = \nconst{normal}\\

                \{ \tabrupt \} &
                \text{otherwise}\\
              \end{array}
            \right.
          }\\

          \aseme{\tname \; \kwcl \cdots \kwcr}(\aenv) &=& \{ \tname \}\\

          \aseme{\expr: \ty}(\aenv) &=& \{ \ty' \subtype \ty \mid \ty' \in
          \aseme{\expr}(\aenv) \}\\

          \aseme{\refer \kwexists}(\aenv) &=& \{ \ty \neq \tabsent \mid \ty \in
          \aseme{\expr}(\aenv) \}\\

          \aseme{\refer}(\aenv) &=& \asemr{\refer}(\aenv) \setminus \{ \tabsent \}\\

          \aseme{\uop \; \expr}(\aenv) &=& \{ \abs{\uop} \ty \mid \ty \in
          \aseme{\expr}(\aenv) \}\\

          \aseme{\expr_0 \bop \expr_1}(\aenv) &=& \{ \ty_0 \abs{\bop} \ty_1 \mid
          \ty_0 \in \aseme{\expr_0}(\aenv) \wedge \ty_1 \in
          \aseme{\expr_1}(\aenv) \}\\
        \end{array}
      $}
      \caption{Expressions with $\aseme{\expr}: \aenvset \rightarrow \atyset$.}
    \end{subfigure}
  \end{minipage}
  \caption{Abstract semantics of the modified $\ires$ for a program $\prog =
  (\getfunc, \getinst, \getnext)$.}
  \vspace*{-1.5em}
  \label{fig:abs-sem}
\end{figure*}

@@

Based on abstract states, we define the abstract semantics of instructions,
references, and expressions in the modified $\ires$ described in
Figure~\ref{fig:abs-sem}.  To prevent the explosion of the number of type-sensitive
views, we upcast the type of each argument before function calls with the
following function:
\begin{figure}[H]
  \centering
  \vspace{-0.5em}
  \resizebox{0.8\columnwidth}{!}{$
    \upcast(\ty) = \left\{
      \begin{array}{ll}
        \tnormal(\upcast(\ty')) & \text{if} \; \ty = \tnormal(\ty')\\
        \clist{\upcast(\ty')} & \text{if} \; \ty = \clist{\ty'}\\
        \tstr & \text{if} \; \ty = \str\\
        \tbool & \text{if} \; \ty = \bool\\
        \ty & \text{otherwise}\\
      \end{array}
    \right.
  $}
  \vspace{-0.5em}
\end{figure} \noindent
and $\upcasts$ denotes a point-wise extension of $\upcast$ for type sequences.
For branches and assertions, we apply the $\pass$ function to prevent the
infeasible control flows defined as follows:

\begin{figure}[H]
  \centering
  \vspace{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \pass(\expr, \bool)(\aenv) = \left\{
      \begin{array}{ll}
        \refine(\expr, \bool)(\aenv)
        & \text{if} \; \{ \true \} \order \aseme{\expr}(\aenv)\\

        \emp
        & \text{otherwise}\\
      \end{array}
    \right.
  $}
  \vspace{-0.5em}
\end{figure} \noindent
where $\refine$ is a function that prunes out infeasible parts in abstract
environments based on their conditions to increase analysis precision.  We will
explain its detail in Section~\ref{sec:refine}.  Then, we define the abstract
semantics $\asem{\prog}$ of a program $\prog = (\getfunc, \getinst, \getnext)$
as the least fixpoint of abstract transfer $\atransfer: \adom \rightarrow
\adom$:
\begin{figure}[H]
  \centering
  \vspace{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}l}
      \asem{\prog} &=& \lim_{n \rightarrow \infty}(\atransfer)^n(\iaelem)\\
      \atransfer(\aelem) &=& \aelem \join \left(
        \bigjoin_{(\lab, \tys) \in \domain{\rmap}}
        \asemi{\getinst(\lab)}(\lab, \tys)(\aelem)
      \right)\\
    \end{array}
  $}
  \vspace{-0.5em}
\end{figure} \noindent
where $\aelem = (\rmap, \_)$ and $\iaelem$ denotes the initial abstract state.
As described in Section~\ref{sec:overview}, $\iaelem$ contains entry points of
all syntax-directed algorithms having no additional parameters and built-in
algorithms with appropriate abstract environments.  For syntax-directed
algorithms, we construct its abstract environment containing the variable
$\this$ with a production type and other variables for non-terminals.  For
example, the upper-left syntax-directed algorithm in Figure~\ref{fig:example} is
initialized with the following abstract environment:
\begin{figure}[H]
  \centering
  \vspace{-0.5em}
  \resizebox{\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}l}
      \this &\mapsto& \{ \code{AssignmentExpression} \},\\
      \code{LeftHandSideExpression} &\mapsto& \{ \code{LeftHandSideExpression} \},\\
      \code{AssignmentExpression} &\mapsto& \{ \code{AssignmentExpression} \}
    \end{array}
  $}
  \vspace{-0.5em}
\end{figure} \noindent
For built-in algorithms, we assign pre-defined variables $\this$, $\args$, and
$\NewTarget$ with corresponding types and parameters with $\tjs$ types.  For
example, the following abstract environment is for the built-in
algorithm $\jscode{Math.round}$ at the upper-right in Figure~\ref{fig:example}:
\begin{figure}[H]
  \centering
  \vspace{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}lr@{~}c@{~}l}
      \this &\mapsto& \{ \tjs \}, &
      \args &\mapsto& \{ \clist{\tjs} \},\\
      \NewTarget &\mapsto& \{ \code{Object}, \undefval \}, &
      \x &\mapsto& \tjs\\
    \end{array}
  $}
  \vspace{-0.5em}
\end{figure}
